Evaluating the Performance and Energy Efficiency of COSMO-ART, a Fully Online Coupled Model System with Numerical Weather Forecast and Chemical Transport Models

### 2014-04-01 trunk/paper_enahpc14/Sec1:COSMO-ART.tex
adds a few updates

\subsection{Model description}
\label{subsec:1.1}
COSMO-ART  (ART  stands  for   Aerosols  and  Reactive  Trace  gases),
developed  at  KIT  Karlsruhe  \citep{Vogel-2009}, is  a  regional  to
continental scale model coupled  online to the COSMO numerical weather
prediction  (NWP) and  climate  model \citep{Baldauf-2011}.   Physical
processes  like  transport,  turbulent  diffusion,  and  dry  and  wet
deposition  are  treated  together  with  photochemistry  and  aerosol
dynamics using  the modal  approach.  Aerosols dynamics  are simulated
with the modal aerosol module MADE \citep{Ackermann-1998}, improved by
explicit  treatment of  soot aging  through condensation  of inorganic
\citep{Riemer-2003}   and   organic   substances   (MADEsoot),   which
represents the  aerosol population using  eleven overlapping lognormal
modes.   Five  modes  represent  sub-micron  particles  consisting  of
sulphate,  ammonium,  nitrate,   organic  compounds,  water  and  soot
\citep{Riemer-2004}  in a  range  of mixing  state.   These modes  are
coupled with  the gas  phase by condensation  and nucleation,  and are
strongly influenced by anthropogenic emissions of gases and particles.
Sea    salt     \citep{Lundgren-2013},    mineral    dust    particles
\citep{Vogel-2006, Stanelle-2010}  and rest anthropogenic  species are
treated by  seven additional modes. For each  mode, mass contributions
and total  number concentration  are prognostic quantities,  while the
standard deviation is fixed.

>>>>

\subsection{Model description}
\label{subsec:1.1}
COSMO-ART  (ART  stands  for   Aerosols  and  Reactive  Trace  gases),
developed  at  KIT  Karlsruhe  \citep{Vogel-2009}, is  a  regional  to
continental scale model coupled  online to the COSMO numerical weather
prediction  (NWP) and  climate  model \citep{Baldauf-2011}.   Physical
processes  like  transport,  turbulent  diffusion,  and  dry  and  wet
deposition  are  treated  together  with  photochemistry  and  aerosol
dynamics  using the  modal approach.

Aerosols dynamics  are simulated  with the modal  aerosol microphysics
module  MADE \citep{Ackermann-1998}, expanded  in MADEsoot  to feature
explicit  treatment of  soot aging  through condensation  of inorganic
\citep{Riemer-2003} and organic substances. MADE describes the aerosol
population  through  five   modes  representing  sub-micron  particles
consisting of sulphate, ammonium, nitrate, particulate organic matter,
water and soot  \citep{Riemer-2004} in a range of  mixing state. These
modes are coupled  with the gas phase by  condensation and nucleation,
and are  strongly influenced by  anthropogenic emissions of  gases and
particles. MADEsoot  describes the sub-micrometer  aerosol population,
composed  of sea  salt \citep{Lundgren-2013},  mineral  dust particles
\citep{Vogel-2006,  Stanelle-2010},   by  means  of   six  interacting
log-normal modes.  For each  mode, mass contributions and total number
concentration are prognostic  quantities, while the standard deviation
is fixed.

### 2014-04-01 trunk/paper_enahpc14/paper.tex
adds a few updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
\documentclass[twocolumn]{svjour3}           % twocolumn
%
\smartqed % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[round, sort&compress, numbers]{natbib}

\usepackage{multirow}
\usepackage{graphicx}
\usepackage{units}
\usepackage{color}

\usepackage{xspace}
\newcommand{\pmlib}{\texttt{pmlib}\xspace}

%
\DeclareRobustCommand\IPCClongname{}
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Evaluating the Performance  and Energy Efficiency of COSMO-ART,
  a Fully Online Coupled Model System with Numerical Weather
    Forecast and Chemical Transport Models }
 % \thanks{This work was supported 
 %   by the EU Project FP7 318793 ``EXA2GREEN''}}
%\subtitle{Write here subtitle}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
	%\\ Tel: +41  (0) 91  610 8216
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        %\and Dr.  William Sawyer 
        %\at Swiss National Supercomputing Centre  (CSCS)
        %\\ CH-6900 Lugano, Switzerland
        %\\ Tel:  +41 (0) 91 610 8229  
        %\\ \email{william.sawyer@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg
	\\ D-22.527 Hamburg, Germany
        %\\ Tel: +49 (0) 40 460094-404
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and C.~Malossi  
        \at IBM Zurich Research Laboratory 
        \\  CH-8000 Zurich, Switzerland
        %\\  Tel: +41 (0) 44 724 8616
        \\ \email{ACM@zurich.ibm.com}}

@@

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  Weather  Service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is  analysed by a  profiling and tracing study  to determine
  hotspots  and  identify   critical  paths.   Moreover,  we  describe
  measurement devices and energy-aware techniques employed to evaluate
  the  energy  footprint of  the  considered  application  and to  get
  detailed  insights about  power bottlenecks.   Our motivation  is to
  improve  corresponding  code sections  to  sustain high  performance
  while  minimizing energy-to-solution.   This preliminary  study sets
  the basis of subsequent  considerations to tackle challenges related
  to energy  efficient high performance computing in  the framework of
  the Exa2Green project (\url{http://exa2green.eu/}).

@@

\section{Power measurement systems}
\label{sec:3}
\input{power-measurement.tex}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
\documentclass[twocolumn]{svjour3}
%
\smartqed
%
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[round, sort&compress, numbers]{natbib}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{units}
\usepackage{color}
\usepackage{xspace}
%
\DeclareRobustCommand\IPCClongname{}
%
% please place your own definitions here and don't use \def but \newcommand{}{}
\newcommand{\pmlib}{\texttt{pmlib}\xspace}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Evaluating the Performance  and Energy Efficiency of COSMO-ART,
  a Fully Online Coupled Model System with Numerical Weather
    Forecast and Chemical Transport Models }
% \thanks{This work was supported by the EU Project FP7 318793 ``EXA2GREEN''}}
%\subtitle{Write here subtitle}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
	%\\ Tel: +41  (0) 91  610 8216
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg
	\\ D-22.527 Hamburg, Germany
        %\\ Tel: +49 (0) 40 460094-404
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and C.~Malossi  
        \at IBM Zurich Research Laboratory 
        \\  CH-8000 Zurich, Switzerland
        %\\  Tel: +41 (0) 44 724 8616
        \\ \email{ACM@zurich.ibm.com}}

@@

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  Weather  Service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is  analysed by a  profiling and tracing study  to determine
  hotspots  and  identify   critical  paths.   Moreover,  we  describe
  measurement devices and energy-aware techniques employed to evaluate
  the  energy  footprint of  the  considered  application  and to  get
  detailed  insights about  power bottlenecks.   Our motivation  is to
  improve  corresponding  code sections  to  sustain high  performance
  while minimizing energy-to-solution.  This preliminary work sets the
  basis for subsequent studies  to tackle challenges related to energy
  efficient  high  performance  computing  in  the  framework  of  the
  Exa2Green project (\url{http://exa2green.eu/}).

@@

\section{Power measurement systems}
\label{sec:3}
\input{Sec3:Power-measurement.tex}
